---
title: "R package development"
author: "Jae Yeon Kim"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

> Anything that can be automated should be automated. Do as little as possible by hand. Do as much as possible with functions. 
- Hadley Wickam

# Why developing R packages? 

1. Reuse your code 
2. Automate your workflow 
3. Help others (be part of an open source development community)

```{r}
# Load packages 
pacman::p_load(devtools, # dev tools 
               usethis, # workflow 
               roxygen2, # documentation 
               testthat, # testing 
               here) # reproducibility 

```

# Workflow 

1. Write code in `\R`
2. Document code in `\man` (automated by `roxygen2` package)
  - `devtools::document()` 
3. Check dependencies in `NAMESPACE`
  - `devtools::update()` updates the documentation 
  - `devtools::check()` to see whether your package is ready to be submitted to CRAN
4. Build a package 
  - `devtools::build()` build a package from sources (a binary version)
5. (Optional) Test (`devtools::test()`), teach in `\vignettes`, and add data in `\data`
6. Distribute the package either via CRAN or GitHub  

# Developing an R package 

The 4 required components are necessary to build and distribute a minimally viable R package. The other steps are optional.

## Required Components

- Package 
  - `\R`: R functions 
  - `\man`: function documentations 
  - DESCRIPTION: provides meta data about the package (e.g., author)
  - LICENSE
    - GNU, MIT, etc.
  - NAMESPACE: package dependencies (to make your package self-contained)
  - README (optional)
  
1. Setup (**DESCRIPTION**)

```{r eval = FALSE}
# This function creates DESCRIPTION file 
usethis::create_package(here("mypkg"))

# Initialize git repo 
usethis::use_git()

# License the package 
# You can use the MIT license by typing devtools::use_mit_license("author name"). The function produces MIT license related files (LICENSE, LICENSE.md).
use_mit_license("Jae Yeon Kim")

# Add README (optional)
# Makes the package more use-friendly 
use_readme_md()

# Add news (optional) 
# Helps track changes 
use_news_md() 
```

2. Write code (**R**)

```{r eval = FALSE}
usethis::use_r("rbind_mutate")
```
```{r eval = FALSE}
#' Add two numbers
#'
#' @param x A number
#' @param y A number
#' @return The sum of x and y 
#' @import package function 
#' @export

add <- function(x, y){
  
  x + y
  
}
```

3. Document (**man**)

```{r eval = FALSE}
# Document 
# The function creates documentation related files (NAMESPACE, function_name.rd)
devtools::document()

# Load all 
devtools::load_all()

# Check 
devtools::check()
```

4. Organize (**NAMESPACE**)

```{r eval = FALSE}
usethis::use_package("dplyr")
```
## Optional Components  

1. Test (**test**)

```{r eval = FALSE}
usethis::use_testthat()

usethis::use_test("rbind_mutate")
```

3. Add data (**data**)

```{r eval = FALSE}

x <- "Jae"
y <- "Sun"
z <- "Jane"

usethis::use_data(x, y, z, overwrite = TRUE)

```
4. Teach (**vignetts**)

```{r eval = FALSE}
usethis::use_vignette("rbind_mutate")
```

```r
title: "Vignette title"
author: "Vignette author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: blah blah
``` 

- You can build a package website using `pkgdown`

```{r eval=FALSE}

# install.packages("pkgdown")
usethis::use_pkgdown()
pkgdown::build_site()

```

- A package site includes information on METADATA, Function references, Articles, News, etc. 

# Building an R package 

- CMD (in the terminal)

You can run R commands in the terminal using R CMD.

```sh
R CMD build mypkg 
R CMD INSTALL mypkg 
```

- devtools 
```{r eval=FALSE}
# Build 
devtools::build()

# Install 
devtools::install()
```

# Distributing an R package 

```{r eval = FALSE}

# Version update 
usethis::use_version()

# Spell check
usethis::use_spell_check()
```

1. [CRAN (The Comprehensive R Archive Network)](https://cran.r-project.org/)
  - R package submission should comply with [the CRAN Repository Policy](https://cran.r-project.org/)
2. GitHub 
  - Push everything to the Git repository (you can do it using command-line interface or RStudio).

```sh 
git add . 
git commit -m "first push"
git push 
```
  - Don't forget that your repository should be `public`.
  - I highly recommend connecting GitHub with SSH. For more information, visit [this link](https://docs.github.com/en/github/authenticating-to-github/connecting-to-github-with-ssh).
  
```sh
git remote set-url origin git@github.com:user/repo 
```