---
title: "R package development"
author: "Jae Yeon Kim"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

> Anything that can be automated should be automated. Do as little as possible by hand. Do as much as possible with functions. 
- Hadley Wickam

# Why developing R packages? 

1. Reuse your code 
2. Automate your workflow 
3. Help others (be part of an open source development community)

# Workflow 

1. Write code 
2. Document code 
3. Check dependencies (NAMESPACE)
4. Build a package 
5. Distribute the package 

# Developing an R package 

The 4 required components are necessary to build and distribute a minimally viable R package. The other steps are optional.

```{r}
# Load packages 
pacman::p_load(devtools, # dev tools 
               usethis, # workflow 
               roxygen2, # documentation 
               testthat) # testing 

# Create a package directory 
path <- file.path(tempdir(), "mypkg")
create_package(path)

# Initialize Git repo 
use_git()

# This creates an associated GitHub repo 
#use_github() 

# Add License 
use_mit_license("Jae Yeon Kim")
```

## 4 Required Components

1. Setup (**DESCRIPTION**)

2. Write code (**R**)

3. Document (**man**)

4. Organize (**NAMESPACE**)

## 3 Optional Components  

1. Test (**test**)

3. Add data (**data**)

4. Teach (**vignetts**)

# Building an R package 

# Distributing an R package 

```{r}
available.packages("")
```

1. [CRAN (The Comprehensive R Archive Network)](https://cran.r-project.org/)
  - R package submission should comply with [the CRAN Repository Policy](https://cran.r-project.org/)
2. GitHub 
  - Push everything to the Git repository (you can do it using command-line interface or RStudio).

```sh 
git add . 
git commit -m "first push"
git push 
```
  - Don't forget that your repository should be `public`.
  - I highly recommend connecting GitHub with SSH. For more information, visit [this link](https://docs.github.com/en/github/authenticating-to-github/connecting-to-github-with-ssh).
  
```sh
git remote set-url origin git@github.com:user/repo 
```